---
title: "GM: Size Class"
bibliography: [book.bib]
---

Two `size classes` (large/small) were used to assess whether the morphology of elliptical bifaces from Millsap Cache and Jowell Farm remains morphologically stable/consistent through reduction.

## Load packages + data

```{r load.packages, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# load required analysis packages
library(here)
library(StereoMorph)
library(geomorph)
library(tidyverse)
library(wesanderson)
library(ggplot2)

# read data and define number of sLMs ----
shapes <- readShapes("shapes")
shapesGM <- readland.shapes(shapes, nCurvePts = c(15, 20, 35))

# read qualitative data
qdata <- read.csv("qdata.merged.csv", 
                  header = TRUE, 
                  row.names = 1)
```

## Generalised Procrustes Analysis

Landmark data were aligned to a global coordinate system [@RN8477;@RN7502;@RN11622;@RN11623;@RN11563], achieved through generalised Procrustes superimposition [@RN11138;@RN478;@RN1646] in R using the `geomorph` and `RRPP` packages [@RN1655;@RN11775;@RN11530;@RN1774;@RN9565]. Procrustes superimposition translates and rotates the coordinate data to allow for comparisons among objects, while also scaling each biface using unit-centroid size---the square root of the sum of squared distances from each landmark to the specimen’s centroid [@RN11139;@RN11140;@RN11564;@RN478]. The `geomorph` package uses a partial Procrustes superimposition that projects the aligned specimens into tangent space subsequent to alignment in preparation for the use of multivariate methods that assume linear space [@RN11141;@RN11142;@RN1646;@RN11563]. 

```{r outlines, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# gpa ----
Y.gpa <- gpagen(shapesGM, print.progress = FALSE)

# geomorph data frame ----
gdf <- geomorph.data.frame(shape = Y.gpa$coords, 
                           size = Y.gpa$Csize,
                           site = qdata$site,
                           size.class = qdata$size,
                           merged = qdata$merged)


# add centroid size to qdata
qdata$csz <- Y.gpa$Csize
```

## Principal Components Analysis

Principal components analysis [@RN1746] was used to visualise shape variation among the elliptical bifaces, and the scatterplot represents the dispersion of shapes in tangent space [@RN8633;@RN5616;@RN11143;@RN7550]. Shape ranges described by each principal axis are commonly visualized using thin-plate spline warping of a reference image [@RN1731;@RN479].

```{r pca.plot, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# principal components analysis ----
pca<-gm.prcomp(Y.gpa$coords)
summary(pca)

# set plot parameters
pch.gps <- c(1:4)[as.factor(qdata$merged)]
col.gps <- wes_palette("Moonrise2")[as.factor(qdata$merged)]
col.hull <- c("#CCC591", "#29211F", "#798E87", "#C27D38")

# plot pca by merged
pc.plot <- plot(pca, 
                asp = 1,
                pch = pch.gps,
                col = col.gps)
shapeHulls(pc.plot, 
           groups = qdata$merged,
           group.cols = col.hull)
```

### Minima/maxima of PC1/2 with warp grids

```{r min.max, echo=TRUE, out.width = "100%", dpi = 300, warning=FALSE}
# plot x/y maxima/minima
## x - minima
mean.shape <- mshape(Y.gpa$coords)
plotRefToTarget(pca$shapes$shapes.comp1$min, 
                mean.shape)

## x - maxima
plotRefToTarget(pca$shapes$shapes.comp1$max, 
                mean.shape)

## y - minima
plotRefToTarget(pca$shapes$shapes.comp2$min, 
                mean.shape)

## y - maxima
plotRefToTarget(pca$shapes$shapes.comp2$max, 
                mean.shape)
```

## Procrustes ANOVA

To assess whether shape and size differ by `size class`, Procrustes ANOVAs [@RN1749] were run that enlist effect-sizes (zscores) computed as standard deviates of the generated sampling distributions [@RN1756]. A residual randomization permutation procedure (RRPP; n = 10,000 permutations) was used for all Procrustes ANOVAs [@RN1655;@RN11775], which has higher statistical power and a greater ability to identify patterns in the data should they be present [@RN1719].

```{r anova, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# ANOVA ----
# shape
fit.sh <- procD.lm(shape ~ merged,
                   data = gdf,
                   print.progress = FALSE,
                   iter = 9999)

# shape as a function of size class?
anova(fit.sh)

# size
fit.sz <- procD.lm(size ~ merged,
                   data = gdf,
                   print.progress = FALSE,
                   iter = 9999)

# size as a function of size class?
anova(fit.sz)

# pairwise comparison of LS means = which differ?
pw.sz <- pairwise(fit.sz,
                  groups = qdata$merged)
summary(pw.sz, 
        confidence = 0.95, 
        test.type = "dist")
```

## Morphological Disparity

An analysis of morphological disparity was used [@RN11107;@RN7041;@RN5694] to identify potential differences in morphological diversity between the size classes.

```{r mdisp, out.width = "100%", dpi = 300, echo=TRUE, warning=FALSE}
# morphological disparity ----
# do any of the groups display greater shape variation among 
# individuals relative to the other group?
morphol.disparity(shape ~ merged,
                  groups = qdata$merged, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)

# do any of the groups display greater size variation among 
# individuals relative to the other group?
morphol.disparity(size ~ merged,
                  groups = qdata$merged, 
                  data = gdf, 
                  print.progress = FALSE, 
                  iter = 9999)
```
